---
title: "SQL"
description: 
  Using SQL to analyze a database
author: Veronica De Leon
date: November 26, 2024
format:
  html:
    code-fold: true
    code-summary: "Show the code"
message: false
warning: false
---

### About the Database

[The link to database](https://www.science.smith.edu/wai-database/)

Using SQL, I queried the Wideband Acoustic Immittance (WAI) Database hosted by Smith College. WAI measurements are made to hopefully provide noninvasive auditory diagnostic tools for everyone. The WAI database hosts WAI measures on ears from all ages(infants, NICU infant, child, adults).

```{r}
library(tidyverse)
```

```{r}
library(RMariaDB)
library(dplyr)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")


# collect(Measurements)
```

### Duplicating Figures

For this analysis, I duplicated figure 1 that's in "An online wideband acoustic immittance (WAI) database and corresponding website" (Voss 2020). The figure shows a line graph of frequency versus mean absorption of the 12 studies found in WAI database as of July 1, 2019. Its legend contains the authors, year, number of individual ears, and equipment used.

```{sql}
#| connection: con_wai
#| output.var: "figure"
SELECT Frequency, Identifier, LOG10(Frequency) AS scaled_freq, AVG(Absorbance) AS avg_absorb
FROM Measurements
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Frequency

```

```{sql}
#| connection: con_wai
#| output.var: "information"
SELECT p.Identifier, p.AuthorsShortList, p.Year, COUNT(DISTINCT SubjectNumber, Ear) AS distinct_ears
FROM PI_Info AS p
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency > 200 AND Frequency < 8000
GROUP BY Identifier, Instrument;
```

```{sql}
#| connection: con_wai
#| output.var: "strings"
SELECT p.Identifier, p.AuthorsShortList, p.Year, AVG(Absorbance) AS avg_absorb, m.Frequency, LOG10(Frequency) AS scaled_freq, CONCAT(p.AuthorsShortList, "(", p.Year, ") N =", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS key_info
FROM PI_Info AS p
LEFT JOIN Measurements AS m ON p.Identifier = m.Identifier
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND m.Frequency > 200 AND m.Frequency < 8000 
GROUP BY Identifier, Instrument, Frequency
HAVING avg_absorb > 0;
```

```{r}
strings |>
  group_by(Identifier) |>
  ggplot(aes(x = Frequency, y = avg_absorb, color = key_info)) +
  geom_line() +
  ylim(0, 1) +
  xlim(200, 8000) +
  labs(x = "Frequency (Hz)",
       y = "Mean Abosrbance",
       title = "Mean absorbance from each publication in WAI database",
       color = "") +
  theme_bw() +
  scale_x_log10() 
```

[Link to the figure I'm duplicating](https://www.ncbi.nlm.nih.gov/core/lw/2.0/html/tileshop_pmc/tileshop_pmc_inline.html?title=Click%20on%20image%20to%20zoom&p=PMC3&id=7093226_nihms-1535071-f0001.jpg)

To duplicate the figure using SQL, I first had to pull 'Measurements' since the table had the absorbance, frequency, and instrument. Through Measurements, I created a table with columns that include the frequency and average absorbance since those are the two main variables being compared in the graph. Because I needed the author names and years of each study, I needed to pull 'PI_Info'. Since all the information I needed are still split into two tables, I used `JOIN`, and joined 'Measurements' and 'PI_Info' through 'Identifier' (since they both had that as a column) to combine the columns I needed and create a table with all the information needed. Once I had the table, I used ggplot to produce a figure that's similar to Figure 1 from Voss 2020. If we compare them, we can see that the graphs are very similar.

### Analysis of Individual Study

I chose one of the 12 studies that included different races, sex, and ethnicity. A limitation present in this database is that many of the studies don't have participants of different races, sex, or ethnicity. However, after taking the time to look at each study, I found that the study (Shaver 2013) included different races, sex, and ethnicity. So, this analysis will compare the frequency versus mean absorption across different races from the study (Shaver 2013).

```{sql}
#| connection: con_wai
#| output.var: "figure2"
SELECT Frequency, Identifier, LOG10(Frequency) AS scaled_freq, AVG(Absorbance) AS average_absorb
FROM Measurements
WHERE Identifier IN ("Shaver_2013")
GROUP BY Identifier, Frequency

```

```{sql}
#| connection: con_wai
#| output.var: "plot2"
SELECT s.Identifier, s.Race, AVG(Absorbance) AS average_absorb, m.Frequency, LOG10(Frequency) AS scaled_freq
FROM Subjects AS s
LEFT JOIN Measurements AS m ON m.Identifier = s.Identifier
WHERE s.Identifier IN ("Shaver_2013")
GROUP BY Identifier, Frequency, Race

```

```{r}
plot2 |>
  ggplot(aes(x = Frequency, y = average_absorb)) +
  geom_line() +
  facet_wrap(~Race) +
  labs(title = "Frequency Versus Mean Absorption Across Different Races",
       x = "Frequency (Hz)",
       y = "Mean Absorption")
```

For this analysis, I used a similar process as the one I used to duplicate Figure 1 (Voss 2020). I used SQL to pull 'Measurements' because it has the data for frequency and absorption. Then I had to pull 'Subjects' because it has the data for different races, sex, and ethnicity. To combine the needed information from both tables, I joined Measurements and Subjects through 'Identifier' and created a new table with just the author (Shaver), frequency, average absorption, race, sex, and ethnicity. Using the new table and ggplot, I created line graphs based on the race.

As seen from the plot above, all of the races included in the study have the same line. This might be because of how the study was conducted and how they counted the ears and measured the absorption. But if we simply focus on the different graphs, across the the different races, the average absorption seems to increase when frequency goes up to 3000 Hz but starts to decrease after that.
