---
title: "Text Analysis"
description: 
  Code and Visuals Using String Functions and Regular Expressions
author: Veronica De Leon
date: September 18, 2024
format:
  html:
    code-fold: true
    code-summary: "Show the code"
message: false
warning: false
---

```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
```

[The link to Netflix data is here](https://www.kaggle.com/datasets/shivamb/netflix-shows)

```{r}
netflix <- read.csv("../veronica-2222.github.io/netflix_titles.csv")
```

### General Netflix Data

```{r}
netflix |>
  select(release_year, date_added) |>
  mutate(added_year = str_extract(date_added, "(?<=, ).+(?=$)")) |>
  filter(!is.na(release_year), !is.na(added_year)) |>
  ggplot(aes(x = release_year, y = added_year)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Year Show/Movie was Released",
    y = "Year Show/Movie was Added",
    title = "Analysis of Year Show/Movie was Released\n\ versus Year Added on Netflix"
  ) +
theme_minimal()
```

The graph displays the difference between the year a show or movie was released versus the year it was added on Netflix. The opaqueness of the points suggests when majority of the shows and movies were added on Netflix. In this case it seems as if majority of the shows and movies were added in 2016 and onward.

```{r}
netflix|>
  select(rating, listed_in) |>
  mutate(tv_rating = str_extract(rating, "(?<=-).+(?=$)")) |>
  mutate(first_listed = str_extract(listed_in, "\\w+")) |>
  filter(first_listed %in% c("Action", "Crime", "Dramas", "Comedies")) |>
  filter(tv_rating %in% c("13", "14", "G", "PG", "MA")) |>
  mutate (level_rating = fct_relevel(tv_rating, c("13", "14", "G", "PG", "MA"))) |>
  ggplot(aes(x = level_rating, fill = first_listed)) +
  geom_bar(position = "fill") +
  labs(
    title = "Number of Specific TV Ratings",
    subtitle = "Categorized by Genres",
    x = "TV-__",
    y = "Proportion of Shows",
    fill = "Genre",
  ) +
  theme_classic()
```

The bar reveals the proportion of shows' first genre they are listed under and their TV-rating. It can reveal which genres were mostly rated TV-14, PG, etc. For example, one of the major points this bar reveals is that majority of the TV-show rated as G are Dramas. The lack of blue suggests that there are few TV shows that have 'Crime' as the first genre they are listed under. And many of the TV-14 and PG shows are listed under 'Comedies' and 'Dramas'.

### 

### Narrowed Netflix Data

##### Comedies and Christmas Eve

```{r}
netflix |>
  select(type, duration, title, date_added, listed_in) |>
  group_by(type) |>
  filter(str_detect(type, "TV Show")) |>
  filter(str_detect(date_added , "2021")) |>
  filter(str_detect(listed_in, "Comedies")) |>
  mutate(season = str_sub(duration, 1, 1)) |> 
  summarize(type, title, season) |>
  ggplot(aes(y = season)) +
  geom_bar() +
  labs(
    title = "Season and Show Count of Comedy TV Shows Added \n\ on Netflix in 2021",
    x = "Number of Shows",
    y = "Number of Seasons"
  ) +
  theme_classic()

```

The bar graph reveals the frequency of comedy shows added on Netflix in 2021 to have a certain number of seasons. It is evident that majority of the comedy shows only have 1 season.

```{r}

Christmas_eve <- netflix |>
  select(title, type, date_added, listed_in) |>
  mutate(first_listed = str_extract(listed_in, "\\w+")) |>
  mutate(Title_Type=  str_c(title, type, first_listed, sep=': ')) |>
  mutate(Date = str_extract(date_added, ".+(?=,)")) |>
  mutate(Genre = listed_in) |>
  filter(Date == "December 24"| Genre == "TV")|>
    select(Title_Type, Date) 

as_tibble(Christmas_eve)
 
```

The table classifies the titles added on Netflix on December 24 (with varying years) with their type (TV Show or Movie) and the first genre they are listed under.
