---
title: "Text Analysis"
description: 
  Code and Visuals Using String Functions and Regular Expressions
author: Veronica De Leon
date: October 2, 2024
format:
  html:
    code-fold: true
    code-summary: "Show the code"
message: false
warning: false
---

```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
```

### About the Data

[Netflix Movies and TV Shows](https://www.kaggle.com/datasets/shivamb/netflix-shows)

```{r}
netflix <- read.csv("../veronica-2222.github.io/netflix_titles.csv")

netflix |>
  group_by(director) |>
  summarize(count = n()) |>
  arrange(desc(count))
```

This data came from kaggle.com and was uploaded by Shivam Bansal. It is regularly updated, but its last update was 3 years ago. The data covers 8,807 Netflix titles, and contains information like their type (TV show or movie), title, director, cast, country, date added on Netflix, actual release year, rating, and duration (as seen from the tibble above).

### General Netflix Data

After observing the data set, I thought it would be interesting to see if there are possible insights that could be found when comparing the year a show or movie was released and the year they were added on Netflix.

```{r}
#| fig.alt: "The scatter plot is titled: Analysis of Year Show/Movie was Released versus Year Added on Netflix. The x-axis is the year the titles are released and the y-axis is the year the title is added on Netflix."

netflix |>
  select(release_year, date_added) |>
  mutate(added_year = str_extract(date_added, "(?<=, ).+(?=$)")) |> #this only takes in the years from the date added column
  filter(!is.na(release_year), !is.na(added_year)) |> #removes the NAs
  ggplot(aes(x = release_year, y = added_year)) + #creates a scatterplot
  geom_point(alpha = 0.3) +
  labs(
    x = "Year Show/Movie was Released",
    y = "Year Show/Movie was Added",
    title = "Analysis of Year Show/Movie was Released\n\ versus Year Added on Netflix"
  ) +
theme_minimal()
```

The graph displays the difference between the year a show or movie was released versus the year it was added on Netflix. The opaqueness of the points suggests when majority of the shows and movies were added on Netflix. In this case it seems as if majority of the shows and movies were added in 2016 and onward.

Focusing on TV shows, I wondered if the proportion of genres differed across different TVs, like TV-13 vs. TV-MA.

```{r}
#| fig.alt: "The bar graph is titled Number of Specific TV Ratings, with the subheading: categorized by genres. The x-axis is TV- and each bar graph is broken down by 13, 14, G, PG, Ma. The proportions are categorized by color relating to a specific color; pink is for action, green for comedies, blue for crime, and purple for dramas."

netflix|>
  select(rating, listed_in) |>
  mutate(tv_rating = str_extract(rating, "(?<=-).+(?=$)")) |> #only takes in the type of tv-rating, removes "TV-"
  mutate(first_listed = str_extract(listed_in, "\\w+")) |>  #creates a first_listed column with just the first genre they're listed in
  filter(first_listed %in% c("Action", "Crime", "Dramas", "Comedies")) |> #filters the first listed in genres
  filter(tv_rating %in% c("13", "14", "G", "PG", "MA")) |> #filters in the tv ratings
  mutate (level_rating = fct_relevel(tv_rating, c("13", "14", "G", "PG", "MA"))) |> #relevels the tv rating so that they're in this order
  ggplot(aes(x = level_rating, fill = first_listed)) + #creates a bar plot
  geom_bar(position = "fill") +
  labs(
    title = "Number of Specific TV Ratings",
    subtitle = "Categorized by Genres",
    x = "TV-__",
    y = "Proportion of Shows",
    fill = "Genre",
  ) +
  theme_classic()
```

The bar reveals the proportion of shows' first genre they are listed under and their TV-rating. It can reveal which genres were mostly rated TV-14, PG, etc. For example, one of the major points this bar reveals is that majority of the TV-show rated as G are Dramas. The lack of blue suggests that there are few TV shows that have 'Crime' as the first genre they are listed under. And many of the TV-14 and PG shows are listed under 'Comedies' and 'Dramas'.

### 

### Narrowed Netflix Data

To further analyze TV shows, I wanted to see if TV shows listed under Comedies for those added in 2021 had an average amount of seasons.

```{r}
#| fig.alt: "The bar graph is titled: Season and Show count of comedy tv shows added on Netflix in 2021. The x-axis is the number of seasons and the y-axis is the number of shows. The plot shows a graph that's right-skewed."

netflix |>
  select(type, duration, title, date_added, listed_in) |>
  group_by(type) |>
  filter(str_detect(type, "TV Show")) |> #filters so that it only includes the TV shows
  filter(str_detect(date_added , "2021")) |> #filters so that it only includes those added in 2021
  filter(str_detect(listed_in, "Comedies")) |> #filters so that it includes those with "Comedies" in their genre
  mutate(season = str_sub(duration, 1, 1)) |>  #filters out the word "season"
  summarize(type, title, season) |>
  ggplot(aes(x = season)) + #creates bar plot
  geom_bar() +
  labs(
    title = "Season and Show Count of Comedy TV Shows Added \n\ on Netflix in 2021",
    x = "Number of Seasons",
    y = "Number of Shows"
  ) +
  theme_classic()

```

The bar graph reveals the frequency of comedy shows added on Netflix in 2021 to have a certain number of seasons. It is evident that majority of the comedy shows only have 1 season.

#### Christmas Analysis

This specific analysis is a little more personal. I always enjoy Christmas movies, so I thought that I could gather a list of TV shows and movies added on Christmas Eve to expand my list of Christmas media.

```{r}
Christmas_eve <- netflix |>
  select(title, type, date_added, listed_in) |> 
  mutate(first_listed = str_extract(listed_in, "\\w+")) |> #filters so that it only includes the first genre they're listed in
  mutate(Title_Type=  str_c(title, type, first_listed, sep=': ')) |> #creates a string with title: type: genre
  mutate(Date = str_extract(date_added, ".+(?=,)")) |> #removes the year from date_added column
  mutate(Genre = listed_in) |>
  filter(Date == "December 24"| Genre == "TV")|> #filters so that it inludes those added on December 24 and are listed as TV
    select(Title_Type) 

as_tibble(Christmas_eve)
 
```

The table classifies the titles added on Netflix on December 24 (with varying years) with their type (TV Show or Movie) and the first genre they are listed under. However, some of them aren't Christmas movies, like Bridezilla and Hello, Love, Goodbye. So, I decided to dig deeper into this data set.

Instead of only looking at the date added, I thought that if a description had the word "Christmas" in it, then maybe the likelihood of getting a Christmas movie or TV show would be higher.

```{r}
Christmas <- netflix |>
  select(title, type, date_added, listed_in, description) |>
  filter(str_detect(description, "Christmas")) |> #filters so that it only includes titles with Christmas in its description
  mutate(first_listed = str_extract(listed_in, "\\w+")) |> #only includes the first genre they're listed in
  mutate(Date = str_extract(date_added, "\\w{3}")) |> #only includes the first three letters of the month for the date they're added
  mutate(Date = as.factor(Date)) |> #factors Date column
  mutate(Date = fct_relevel(Date, c("Jan", "Feb", "Mar", "Jun", "Jul", "Oct", "Nov", "Dec"))) |> #had to reorder the months 
  mutate(Title_Type =  str_c(title, type, first_listed, sep=': ')) |> #creates a string with title:type:genre
  summarize(Title_Type, Date) 

head(Christmas)
```

The table classifies titles on Netflix with the word "Christmas" in its description with their title, type, first genre listed. The column 'Date' was cleaned up so that it would only include the first three letters of the months. By only including descriptions that have the word "Christmas" in it, I was able to retrieve a greater list of Christmas movies and shows, as seen from the tibble above.

```{r}
#| fig.alt: "The bar graph is titled Count of Movies and TV shows based on month added. The x-axis is the month and the y-axis is the count of the titles. The bars are colored in dark green and the axes titles are colored in dark red."

#creates a bar plot with months as x-axis
Christmas |>
  ggplot(aes(x = Date)) +
  geom_bar(fill = "darkgreen") +
  labs(
    title = "Count of Movies and TV Shows based on Month Added",
    x = "Month", 
    y = "Count"
  ) +
  theme(
    axis.title = element_text(color = "darkred")
  )
```

Since I didn't get a list of Christmas movies after filtering titles added on December 24, I also created this bar graph from the table above it to see when such titles with "Christmas" in its description are added on Netflix. From the bar graph, it appears that majority of them are added on December and November, so maybe Netflix adds them throughout the month of December and not just Christmas Eve. I also found it quite surprising that after observing this graph, there's a couple of titles added in July and July, months I wouldn't expect to see Christmas movies added on Netflix. The titles added in June and July are seen in the table below.

```{r}
July <- Christmas|>
  filter(str_detect(Date, "Jul"))  #creates a table with just titles with Christmas in its description that are added in July

June <- Christmas |>
  filter(str_detect(Date, "Jun")) #creates a table with just titles with Christmas in its description that are added in June


Summer <- rbind(July, June) #creates a table combining the two tables above it
Summer
```
