---
title: "A Simulation Study"
description: 
  Creating a Visual Inference 
author: Veronica De Leon
date: November 8, 2024
format:
  html:
    code-fold: true
    code-summary: "Show the code"
message: false
warning: false
---

```{r}
library(openintro)
library(tidyverse)
library(stringr)
```

### About the Data

[The link to the original data is here](https://openintrostat.github.io/openintro/reference/dream.html)

For this project, I used the 'Dream' data set from the openintro package. The data set came from SurveyUSA, News Poll and was collected on January 27-29, 2012. The poll surveyed the views of 910 people on the DREAM Act (a legislative proposed to give undocumented immigrants temporary residency in the US). Those surveyed were categorized into three ideologies: Liberal, Moderate, and Conservative. Their responses were either No, Not Sure, or Yes.

```{r}
table(dream)
```

From the table, it appears that the there isn't an equal number of people for each ideology. So, the number for each stance isn't a great indicator of whether or not a certain ideology leans more towards a certain stance. Thus, if we were to analyze and compare the information from this data, we would have to look at the proportions.

## About the Simulation

From the observed data of the data set, I noticed that the proportion of Liberals in support of the Dream Act (0.65) is greater than the proportion of Conservatives (0.5) and the proportion of Moderates (0.48). So, I set to test if there is truly a relationship between ideology and whether they are in support of the DREAM Act by comparing null plots and the observed plot.

Under the null, the proportions between ideologies in support of the DREAM Act are equal . So, when we compare the null plots and observed plot, we are checking to see if our observed plot differs from the null plots. If we can identify the observed plot from the null plots, then this gives us evidence to reject the null, and believe that there is a relationship between ideology and their support of the DREAM Act.

## Simulation

The data set was edited so that the rows of the ideology column only included the first letter of each of the ideologies for simplicity. They were also releveled so that Liberal was on the left, Moderate was in the middle, and Conservative was on the right.

```{r}
dream <- dream |>
  mutate(ideology = as.character(ideology)) |>
  mutate(ideology = str_extract(ideology, "\\w")) |> #only takes the first letter of the word
  mutate(ideology = as.factor(ideology)) |>
  mutate(ideology = fct_relevel(ideology, c("L", "M", "C")))  #relevels the ideologies in order L, M, C
```

Here is the "statistic" in a form of a bar graph. This is our observed plot as it uses the information provided by the data set. We can see that the proportion of 'Yes', shaded in blue, is greater for Liberal.

```{r}
#| fig.alt: "A bar graph titled Observed Plot. The x-axis are the ideologies: Liberal, Moderate, and Conservative. The y-axis goes from 0 to 1 to show the proportions for each stance (pink is no, green is not sure, and blue is yes)."

#to know which panel to randomly place the plot in when comparing with the null
set.seed(47)
sample(c(1:20), size = 1)

obs_dream <- dream |>
  group_by(ideology) |>
  mutate(repeated = 18)

obs_dream|>
   ggplot(aes(x = ideology, fill = stance)) +
    geom_bar(position = "fill") +
   labs(x = "Ideology", y = "Proportion of Responses", title = "Observed Plot", fill = "Stance")

```

Then I created a function that conducts a randomization test by reshuffling the ideologies, but keeping the stance the same for each row.

```{r}
#created a function that shuffles ideology column
shuffle <- function(rep){
  dream |>
    mutate(ideology = sample(ideology, replace = FALSE)) |>
    mutate(repeated = rep)
}
```

Here I mapped out 19 data frames (leaving out 18 because that's reserved for the observed) that outputs what we'd expect under the null, or if the likelihood of each ideology to have a certain stance is equal.

```{r}
#to get data that would appear under the null
set.seed(47)
null_dream_one <- map((1:17), shuffle)|>
    list_rbind() 

null_dream_two <- map((19:20), shuffle) |>
  list_rbind() 

```

Finally, I combined all the data frames (the statistic and the 19 null data frames) and combined all of the bar graphs from the 20 data frames. In the very beginning, I used a sample to generate a number so that I know which panel to place the observed plot in. Through that, the observed plot was placed in Panel #18.

```{r}
#| fig.alt: "A bar graph titled Comparison of Null Plots versus Observed Plot. There are 20 total bar plots. The x-axis are the ideologies: Liberal, Moderate, and Conservative. The y-axis goes from 0 to 1 to show the proportions for each stance (pink is no, green is not sure, and blue is yes)."

#to plot all the graphs
rbind(obs_dream, null_dream_one, null_dream_two) |>
  ggplot(aes(x = ideology, fill = stance)) +
  geom_bar(position = "fill") +
  facet_wrap(~repeated) +
  labs(x = "Ideology", y = "Proportion of Responses", title = "Comparison of Null Plots vs. Observed Plot", fill = "Stance")
```

As we can see from the figure above, most of the shading of blue (which is what we're interested in as it shows the proportion of "Yes") is almost leveled, except for Panel #18. So, the observed plot (Panel #18) does stand out from the rest of the bar graphs as the shading of blue for "L" is greater compared to the others. This figure provides evidence to reject the null hypothesis, which hypothesized that the proportion of "Yes" is equal among the ideologies.

### Summary of Simulation

I began this study by observing the data set and noticing that there is a greater proportion of Liberals who are in support of the DREAM Act. I tested this hypothesis by shuffling the order of the ideologies, via my `shuffle` function, to see if it was common to get the order of ideologies of the original data set. To compare the observation from the 19 reshuffles, I plotted their bar graphs and checked to see if the observed bar graph was identifiable among the bar graphs of the shuffled data frames, and it was.

Because the observed plot is identifiable, we have evidence to reject the null and conclude that there is a difference among the proportion of those in support of the DREAM Act across the three political ideologies. Since the data came from a SurveyUSA poll, the results of this study describes Americans who identify with one of the three political ideologies used in the survey. So based on the study, we can conclude that Americans who identify as liberal are more likely to be in support of the DREAM Act than Americans who identify as moderate or conservative. However, we also need to take into account that the data came from a survey conducted in 2012, which was 12 years ago. Which means that although the study lets us generalize the results to the American population, it is limited to Americans during the 2010s as immigration policies have changed since then, and people affiliated with the three ideologies might have changed their stance on the DREAM Act if they were to be asked today.
